---
title: "BST 210 Project"
date: '2022-12-05'
output: github_document
---

```{r setup, include=FALSE}
library(tidyverse)
library(ipumsr)
library(nnet)
library(VGAM)
library(broom)
library(ResourceSelection)
library(pROC)
library(regclass)
library(splines2)
library(gam)
# library(mice)
ddi <- read_ipums_ddi("nhis_00002.xml")
raw_data <- read_ipums_micro(ddi)

knitr::opts_chunk$set(warning = FALSE)
```

# Data Cleaning

```{r}
data <-
  raw_data %>%
  dplyr:: select(AGE, SEX, RACENEW, INCFAM07ON, HEALTH, MOD10DMIN, MOD10FWK, ANXIETYEV, DEPFREQ, EMOSUPPORT, CVDSYMP, REGION) %>% 
  filter(MOD10DMIN != 000) %>% 
  filter(MOD10FWK != 00) %>% 
  filter(ANXIETYEV != 0) %>% 
  filter(DEPFREQ != 0) %>% 
  filter(EMOSUPPORT != 0) %>% 
  filter(CVDSYMP != 0)
data %>% summary()
```

```{r}
data$AGE[data$AGE == 997] = NA
data$AGE[data$AGE == 999] = NA
summary(data$AGE)
```

```{r}
data$sexcat <- NA
data$sexcat[data$SEX == 1] = "male"
data$sexcat[data$SEX == 2] = "female"
data$sexcat <- as.factor(data$sexcat)
summary(data$sexcat)
```

```{r}
data$racecat <- NA
data$racecat[data$RACENEW == 100] = "white"
data$racecat[data$RACENEW == 200] = "non-white"
data$racecat[data$RACENEW == 300] = "non-white"
data$racecat[data$RACENEW == 400] = "non-white"
data$racecat[data$RACENEW > 400 & data$RACENEW < 900] = "non-white"
data$racecat <- as.factor(data$racecat)
summary(data$racecat)
```

```{r}
data$incomecat <- NA
data$incomecat[data$INCFAM07ON == 11] = "under $49,999"
data$incomecat[data$INCFAM07ON == 12] = "under $49,999"
data$incomecat[data$INCFAM07ON == 22] = "$50,000-$99,999"
data$incomecat[data$INCFAM07ON == 23] = "$50,000-$99,999"
data$incomecat[data$INCFAM07ON == 24] = "over $100,000"
data$incomecat <- factor(data$incomecat, levels = c("under $49,999", "$50,000-$99,999", "over $100,000"))
summary(data$incomecat)
```

```{r}
data$healthcat <- NA
data$healthcat[data$HEALTH == 1] = "good"
data$healthcat[data$HEALTH == 2] = "good"
data$healthcat[data$HEALTH == 3] = "good"
data$healthcat[data$HEALTH == 4] = "fair"
data$healthcat[data$HEALTH == 5] = "poor"
data$healthcat <- factor(data$healthcat, levels = c("good", "fair", "poor"))
summary(data$healthcat)
```

```{r}
data$MOD10DMIN[data$MOD10DMIN %in% c(996, 997, 998, 999)] = NA
summary(data$MOD10DMIN)
```

```{r}
data$anxiety <- NA
data$anxiety[data$ANXIETYEV == 1] = "No"
data$anxiety[data$ANXIETYEV == 2] = "Yes"
data$anxiety <- factor(data$anxiety, levels = c("No", "Yes"))
summary(data$anxiety)
```

```{r}
data$depress <- NA
data$depress[data$DEPFREQ == 1] = "daily"
data$depress[data$DEPFREQ == 2] = "weekly or monthly"
data$depress[data$DEPFREQ == 3] = "weekly or monthly"
data$depress[data$DEPFREQ == 4] = "yearly"
data$depress[data$DEPFREQ == 5] = "never"
data$depress <- factor(data$depress, levels = c("never", "daily", "weekly or monthly", "yearly"))
summary(data$depress)
```

```{r}
data$esupport <- NA
data$esupport[data$EMOSUPPORT == 1] = "always"
data$esupport[data$EMOSUPPORT == 2] = "usually"
data$esupport[data$EMOSUPPORT == 3] = "sometimes-never"
data$esupport[data$EMOSUPPORT == 4] = "sometimes-never"
data$esupport[data$EMOSUPPORT == 5] = "sometimes-never"
data$esupport <- factor(data$esupport, levels = c("always", "usually", "sometimes-never"))
summary(data$esupport)
```

```{r}
data$sympt <- NA
data$sympt[data$CVDSYMP == 1] = "no symptoms"
data$sympt[data$CVDSYMP == 2] = "mild symptoms"
data$sympt[data$CVDSYMP == 3] = "moderate symptoms"
data$sympt[data$CVDSYMP == 4] = "severe symptoms"
data$sympt <- factor(data$sympt, levels = c("no symptoms", "mild symptoms", "moderate symptoms", "severe symptoms"))
summary(data$sympt)
```

```{r}
data$reg <- NA
data$reg[data$REGION == 1] = "northeast"
data$reg[data$REGION == 2] = "midwest"
data$reg[data$REGION == 3] = "south"
data$reg[data$REGION == 4] = "west"
data$reg <- factor(data$reg, levels = c("northeast", "midwest", "south", "west"))
summary(data$reg)
```

```{r}
# data <- data %>% 
#   dplyr::select(12:20, AGE, MOD10DMIN, MOD10FWK)
data %>% 
  summary()
dim(data)[1]*0.05

colnames(data)
```

# Modeling
# Kenny's Analyses

# Kenny's EDA

```{r}
#univariate
hist(data$MOD10DMIN)
# tally(~data$MOD10FWK)

hist(log(data$MOD10DMIN))
hist(data$AGE)
barplot(prop.table(table(data$sympt)))
barplot(prop.table(table(data$sexcat)))
barplot(prop.table(table(data$racecat)))
barplot(prop.table(table(data$incomecat)))
barplot(prop.table(table(data$healthcat)))
ggplot(data, aes(sympt, MOD10DMIN)) +   
  geom_boxplot() +
  labs(x = "Severity of COVID-19 Symptoms", y = "Duration of Moderate Physical Activity")
```

We see that the duration of moderate physical activity is heavily right skewed so we log transformed it.

```{r}
#bivariate
scatter.smooth(data$AGE, log(data$MOD10DMIN))
scatter.smooth(data$AGE, (data$MOD10DMIN))
boxplot(data$MOD10DMIN ~ data$sympt)
boxplot(data$MOD10DMIN ~ data$racecat)
boxplot(data$MOD10DMIN ~ data$incomecat)
boxplot(data$MOD10DMIN ~ data$healthcat)
```


# Fitting the full model
```{r}
lm1 <- lm(log(MOD10DMIN) ~ sexcat + racecat  + healthcat + sympt + AGE
      ,data = data)

summary(lm1)
```

In the full model, we see that after adjusting for the effect of the other variables, none of the COVID symptom levels have a statistically significant effect on the duration of moderate physical activity.

# Effect modifier:
```{r}
# check if anxiety, depression, region or emotional support are effect modifiers of covid symptoms
lm_anxiety <- lm(log(MOD10DMIN) ~ sexcat + racecat  + healthcat + sympt + AGE
          + anxiety*sympt,data = data)
lm_depress <- lm(log(MOD10DMIN) ~ sexcat + racecat  + healthcat + sympt + AGE
          + depress*sympt,data = data)
lm_reg <- lm(log(MOD10DMIN) ~ sexcat + racecat  + healthcat + sympt + AGE
          + reg*sympt,data = data)
lm_esupport <- lm(log(MOD10DMIN) ~ sexcat + racecat + incomecat + healthcat + sympt + AGE
          + esupport*sympt,data = data)
lm_income <- lm(log(MOD10DMIN) ~ sexcat + racecat  + healthcat + sympt + AGE
          + incomecat*sympt,data = data)

summary(lm_anxiety)
summary(lm_depress)
summary(lm_reg)
summary(lm_esupport)
summary(lm_income)
# None are effect modifiers so we will stick with
lm1 <- lm(log(MOD10DMIN) ~ sexcat + racecat + healthcat + sympt + AGE, data = data)

summary(lm1)
```


# Diagnosis and Model Evaluation

```{r}
plot(lm1, 1:6)
hist(lm1$residuals)
# 265, 46, 383

# 12/397
# 0.0302267

# modelling dropping those 
lm1_outliers <- lm(log(MOD10DMIN) ~ sexcat + racecat + healthcat + sympt + AGE, data = 
data[-c(46,265, 383),])

summary(lm1_outliers)
summary(lm1)

# none are above 5 :)
VIF(lm1) 
```
From the residual plots, it seems linearity, normality, and equal variance holds.