---
title: "BST 210 Project"
author: "Yunyang Zhong"
date: '2022-11-03'
output: html_document
---

```{r setup, include=FALSE}
library(tidyverse)
library(ipumsr)
library(nnet)
library(VGAM)
library(broom)
library(ResourceSelection)
library(pROC)
library(regclass)
ddi <- read_ipums_ddi("nhis_00002.xml")
raw_data <- read_ipums_micro(ddi)
data <- read_ipums_micro(ddi)


knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE)
```


```{r}
data$AGE[data$AGE == 997] = NA
data$AGE[data$AGE == 999] = NA
data$AGE[data$AGE == 0] = NA

summary(data$AGE)
```

```{r}
data$sexcat <- NA
data$sexcat[data$SEX == 1] = "male"
data$sexcat[data$SEX == 2] = "female"
data$sexcat <- as.factor(data$sexcat)

summary(data$sexcat)
```

```{r}
data$sexoricat <- NA
data$sexoricat[data$SEXORIEN == 1] = "homosexual"
data$sexoricat[data$SEXORIEN == 2] = "heterosexual"
data$sexoricat[data$SEXORIEN == 3] = "bisexual"
data$sexoricat[data$SEXORIEN == 4] = "other"
data$sexoricat <- factor(data$sexoricat, levels = c("homosexual", "heterosexual", "bisexual", "other"))

summary(data$sexoricat)
```

```{r}
data$racecat <- NA
data$racecat[data$RACENEW == 100] = "white"
data$racecat[data$RACENEW == 200] = "black"
data$racecat[data$RACENEW == 300] = "aian"
data$racecat[data$RACENEW == 400] = "asian"
data$racecat[data$RACENEW > 400 & data$RACENEW < 900] = "other"
data$racecat <- as.factor(data$racecat)

summary(data$racecat)
```

```{r}
data$HOURSWRK[data$HOURSWRK > 95] = NA
data$HOURSWRK[data$HOURSWRK == 0] = NA

summary(data$HOURSWRK)
```

```{r}
data$incomecat <- NA
data$incomecat[data$INCFAM07ON == 11] = "under $34,999"
data$incomecat[data$INCFAM07ON == 12] = "$35,000-$49,999"
data$incomecat[data$INCFAM07ON == 22] = "$50,000-$74,999"
data$incomecat[data$INCFAM07ON == 23] = "$75,000-$99,999"
data$incomecat[data$INCFAM07ON == 24] = "over $100,000"
data$incomecat <- factor(data$incomecat, levels = c("under $34,999", "$35,000-$49,999", "$50,000-$74,999", "$75,000-$99,999", "over $100,000"))

summary(data$incomecat)
```

```{r}
data$healthcat <- NA
data$healthcat[data$HEALTH == 1] = "excellent"
data$healthcat[data$HEALTH == 2] = "very good"
data$healthcat[data$HEALTH == 3] = "good"
data$healthcat[data$HEALTH == 4] = "fair"
data$healthcat[data$HEALTH == 5] = "poor"
data$healthcat <- factor(data$healthcat, levels = c("excellent", "very good", "good", "fair", "poor"))

summary(data$healthcat)
```

```{r}
data$insurcat <- NA
data$insurcat[data$HINOTCOVE == 1] = "no"
data$insurcat[data$HINOTCOVE == 2] = "yes"
data$insurcat <- as.factor(data$insurcat)

summary(data$insurcat)
```

```{r}
data$MOD10DMIN[data$MOD10DMIN %in% c(0, 996, 997, 998, 999)] = NA

summary(data$MOD10DMIN)
```

```{r}
data$MOD10FWK[data$MOD10FWK == 0] = NA
data$MOD10FWK[data$MOD10FWK > 90] = NA

summary(data$MOD10FWK)
```

```{r}
data$anxiety <- NA
data$anxiety[data$ANXIETYEV == 1] = "No"
data$anxiety[data$ANXIETYEV == 2] = "Yes"
data$anxiety <- as.factor(data$anxiety)

summary(data$anxiety)
```

```{r}
data$depress <- NA
data$depress[data$DEPFREQ == 1] = "daily"
data$depress[data$DEPFREQ == 2] = "weekly"
data$depress[data$DEPFREQ == 3] = "monthly"
data$depress[data$DEPFREQ == 4] = "yearly"
data$depress[data$DEPFREQ == 5] = "never"
data$depress <- factor(data$depress, levels = c("daily", "weekly", "monthly", "yearly", "never"))

summary(data$depress)
```

```{r}
data$esupport <- NA
data$esupport[data$EMOSUPPORT == 1] = "always"
data$esupport[data$EMOSUPPORT == 2] = "usually"
data$esupport[data$EMOSUPPORT == 3] = "sometimes"
data$esupport[data$EMOSUPPORT == 4] = "rarely"
data$esupport[data$EMOSUPPORT == 5] = "never"
data$esupport <- factor(data$esupport, levels = c("always", "usually", "sometimes", "rarely", "never"))

summary(data$esupport)
```

```{r}
data$covid <- NA
data$covid[data$CVDTESTRSLT == 1] = "No"
data$covid[data$CVDTESTRSLT == 2] = "Yes"
data$covid <- as.factor(data$covid)

summary(data$covid)
```

```{r}
data$sympt <- NA
data$sympt[data$CVDSYMP == 1] = "no symptoms"
data$sympt[data$CVDSYMP == 2] = "mild symptoms"
data$sympt[data$CVDSYMP == 3] = "moderate symptoms"
data$sympt[data$CVDSYMP == 4] = "severe symptoms"
data$sympt <- factor(data$sympt, levels = c("no symptoms", "mild symptoms", "moderate symptoms", "severe symptoms"))

summary(data$sympt)
```

```{r}
data$reg <- NA
data$reg[data$REGION == 1] = "northeast"
data$reg[data$REGION == 2] = "midwest"
data$reg[data$REGION == 3] = "south"
data$reg[data$REGION == 4] = "west"
data$reg <- factor(data$reg, levels = c("northeast", "midwest", "south", "west"))

summary(data$reg)
```

```{r, results='hold'}
data <- data %>% 
  dplyr::select(32:42, AGE, HOURSWRK, MOD10DMIN, MOD10FWK, HISPETH, EDUC, reg)

data %>% 
  summary()

dim(data)[1]*0.05

dim(data %>% drop_na())[1]
```

> 4 numeric variables: age, total hours worked last week usually, duration of moderate activity 10+ minutues, frequency of moderate activity 10+ minutes

> 2 un-cleaned variables: hispanic ethnicity, educational attainment



```{r}
data %>% 
  ggplot(aes(x = AGE)) +
  geom_density()

data %>% 
  ggplot(aes(x = HOURSWRK)) +
  geom_density()

data %>% 
  ggplot(aes(x = MOD10DMIN)) +
  geom_density()

data %>% 
  ggplot(aes(x = MOD10FWK)) +
  geom_density()
```

```{r}
data %>% 
  filter(anxiety == "Yes") %>% 
  summary()
```


# Kenny's Analyses

```{r}
filtered_data <- data |> 
  dplyr::select(sexcat, racecat, incomecat, healthcat, 
         sympt, AGE, MOD10DMIN, 
         MOD10FWK)

complete_data <- filtered_data |> 
  drop_na()
```

We have 8 variables in our data set with 378 complete cases.
We are interested in modeling the relationship between the severity of covid-19 symptoms and the duration of moderate physical activity. In our model, we are adjusting for sex, race, income, health, age.

# Kenny's EDA

```{r}
#univariate
hist(complete_data$MOD10DMIN)
hist(log(complete_data$MOD10DMIN))
hist(complete_data$AGE)
barplot(prop.table(table(complete_data$sympt)))
barplot(prop.table(table(complete_data$sexcat)))
barplot(prop.table(table(complete_data$racecat)))
barplot(prop.table(table(complete_data$incomecat)))
barplot(prop.table(table(complete_data$healthcat)))
```

We see that the duration of moderate physical activity is heavily right skewed so we log transformed it.

```{r}
#bivariate
scatter.smooth(complete_data$AGE, log(complete_data$MOD10DMIN))
boxplot(complete_data$MOD10DMIN ~ complete_data$sympt)
boxplot(complete_data$MOD10DMIN ~ complete_data$racecat)
boxplot(complete_data$MOD10DMIN ~ complete_data$incomecat)
boxplot(complete_data$MOD10DMIN ~ complete_data$healthcat)

# checking for collinearity between health and covid symptoms
chisq.test(complete_data$healthcat, complete_data$sympt)
```

There does not appear to be a linear relationship between age and log(MOD10DMIN) so we might not want to include it in this model. We will keep it for this hw checkin and do some flexible modeling below.

# Fitting the full model
```{r}
lm1 <- lm(log(MOD10DMIN) ~ .,data = complete_data |> dplyr::select(-MOD10FWK))
# lm1_reg <- lm(log(MOD10DMIN)~ .,data = complete_data)
tidy(lm1)
# summary(lm1)
```

In the full model, we see that after adjusting for the effect of the other variables, none of the COVID symptom levels have a statistically significant effect on the duration of moderate physical activity.

# Diagnosis and Model Evaluation

```{r}
plot(lm1, 1:6)
```
From the residual plots, it seems linearity, normality, and equal variance holds.


# Flexible modeling
```{r}
library(splines2)
library(gam)
gam1 <- gam(log(MOD10DMIN) ~ s(AGE, 4) + .,data = complete_data |> dplyr::select(-MOD10FWK))
anova(lm1, gam1)
```
It seems that it is better to use the GAM model instead of the linear term for Age


# Maria's Poisson

```{r}
hist(complete_data$MOD10FWK)
hist(log(complete_data$MOD10FWK))

```


```{r}
mean(complete_data$MOD10FWK)
var(complete_data$MOD10FWK)


mod_poisson1 <- glm(MOD10FWK ~ . , data = complete_data |> dplyr::select(-MOD10DMIN) , family = poisson())
tidy(mod_poisson1)
deviance(mod_poisson1)/mod_poisson1$df.residual
```

