---
title: "BST 210 Project"
date: '2022-12-05'
output: github_document
---

```{r setup, include=FALSE}
library(tidyverse)
library(ipumsr)
library(nnet)
library(VGAM)
library(broom)
library(ResourceSelection)
library(pROC)
library(regclass)
library(splines2)
library(gam)
library(groupdata2)
library(caret)
library(generalhoslem)
ddi <- read_ipums_ddi("nhis_00002.xml")
data <- read_ipums_micro(ddi)

knitr::opts_chunk$set(warning = FALSE)
```

# Data Cleaning

```{r}
data <-
  data %>%
  dplyr::select(AGE, SEX, RACENEW, INCFAM07ON, HEALTH, MOD10DMIN, ANXIETYEV, DEPFREQ, EMOSUPPORT, CVDSYMP, REGION) %>% 
  filter(MOD10DMIN != 000) %>% 
  filter(ANXIETYEV != 0) %>% 
  filter(DEPFREQ != 0) %>% 
  filter(EMOSUPPORT != 0) %>% 
  filter(CVDSYMP != 0)
```

```{r}
data$AGE[data$AGE == 997] = NA
data$AGE[data$AGE == 999] = NA
```

```{r}
data$sexcat <- NA
data$sexcat[data$SEX == 1] = "male"
data$sexcat[data$SEX == 2] = "female"
data$sexcat <- as.factor(data$sexcat)
```

```{r}
data$racecat <- NA
data$racecat[data$RACENEW == 100] = "white"
data$racecat[data$RACENEW == 200] = "non-white"
data$racecat[data$RACENEW == 300] = "non-white"
data$racecat[data$RACENEW == 400] = "non-white"
data$racecat[data$RACENEW > 400 & data$RACENEW < 900] = "non-white"
data$racecat <- as.factor(data$racecat)
```

```{r}
data$incomecat <- NA
data$incomecat[data$INCFAM07ON == 11] = "under $49,999"
data$incomecat[data$INCFAM07ON == 12] = "under $49,999"
data$incomecat[data$INCFAM07ON == 22] = "$50,000-$99,999"
data$incomecat[data$INCFAM07ON == 23] = "$50,000-$99,999"
data$incomecat[data$INCFAM07ON == 24] = "over $100,000"
data$incomecat <- factor(data$incomecat, levels = c("under $49,999", "$50,000-$99,999", "over $100,000"))
```

```{r}
data$healthcat <- NA
data$healthcat[data$HEALTH == 1] = "good"
data$healthcat[data$HEALTH == 2] = "good"
data$healthcat[data$HEALTH == 3] = "good"
data$healthcat[data$HEALTH == 4] = "fair"
data$healthcat[data$HEALTH == 5] = "poor"
data$healthcat <- factor(data$healthcat, levels = c("good", "fair", "poor"))
```

```{r}
data$MOD10DMIN[data$MOD10DMIN %in% c(996, 997, 998, 999)] = NA
```

```{r}
data$anxiety <- NA
data$anxiety[data$ANXIETYEV == 1] = "No"
data$anxiety[data$ANXIETYEV == 2] = "Yes"
data$anxiety <- factor(data$anxiety, levels = c("No", "Yes"))
```

```{r}
data$depress <- NA
data$depress[data$DEPFREQ == 1] = "daily"
data$depress[data$DEPFREQ == 2] = "weekly or monthly"
data$depress[data$DEPFREQ == 3] = "weekly or monthly"
data$depress[data$DEPFREQ == 4] = "yearly"
data$depress[data$DEPFREQ == 5] = "never"
data$depress <- factor(data$depress, levels = c("never", "daily", "weekly or monthly", "yearly"))
```

```{r}
data$esupport <- NA
data$esupport[data$EMOSUPPORT == 1] = "always"
data$esupport[data$EMOSUPPORT == 2] = "usually"
data$esupport[data$EMOSUPPORT == 3] = "sometimes-never"
data$esupport[data$EMOSUPPORT == 4] = "sometimes-never"
data$esupport[data$EMOSUPPORT == 5] = "sometimes-never"
data$esupport <- factor(data$esupport, levels = c("always", "usually", "sometimes-never"))
```

```{r}
data$sympt <- NA
data$sympt[data$CVDSYMP == 1] = "no symptoms"
data$sympt[data$CVDSYMP == 2] = "mild symptoms"
data$sympt[data$CVDSYMP == 3] = "moderate symptoms"
data$sympt[data$CVDSYMP == 4] = "severe symptoms"
data$sympt <- factor(data$sympt, levels = c("no symptoms", "mild symptoms", "moderate symptoms", "severe symptoms"))
```

```{r}
data$reg <- NA
data$reg[data$REGION == 1] = "northeast"
data$reg[data$REGION == 2] = "midwest"
data$reg[data$REGION == 3] = "south"
data$reg[data$REGION == 4] = "west"
data$reg <- factor(data$reg, levels = c("northeast", "midwest", "south", "west"))
```

```{r}
data <- data %>% 
  dplyr::select(12:20, AGE, MOD10DMIN)

data %>% 
  summary()

# dim(data)[1]*0.05 # 5% missing 22.3

data <- data %>% drop_na()

# dim(data) # 397*11
```

# Data Visualization

```{r}
data %>% 
  ggplot(aes(x=sexcat, fill=anxiety)) +
  geom_bar(position="fill")
data %>% 
  ggplot(aes(x=sexcat, fill=depress)) +
  geom_bar(position="fill")
```

> Female seems to be positively associated with anxiety and depression.

```{r}
data %>% 
  ggplot(aes(x=racecat, fill=anxiety)) +
  geom_bar(position="fill")
data %>% 
  ggplot(aes(x=racecat, fill=depress)) +
  geom_bar(position="fill")
```

> White and other seem to be at higher risk of anxiety and depression.

```{r}
data %>% 
  ggplot(aes(x=incomecat, fill=anxiety)) +
  geom_bar(position="fill")
data %>% 
  ggplot(aes(x=incomecat, fill=depress)) +
  geom_bar(position="fill")
```

> Higher income seems to be a protective factor.

```{r}
data %>% 
  ggplot(aes(x=healthcat, fill=anxiety)) +
  geom_bar(position="fill")
data %>% 
  ggplot(aes(x=healthcat, fill=depress)) +
  geom_bar(position="fill")
```

> Poor health status is highly associated with anxiety and especially depression.

```{r}
data %>% 
  ggplot(aes(x=esupport, fill=anxiety)) +
  geom_bar(position="fill")
data %>% 
  ggplot(aes(x=esupport, fill=depress)) +
  geom_bar(position="fill")
```

> Participants who rarely or sometimes receive emotional support seem to be under higher risk of anxiety and depression.

```{r}
data %>% 
  ggplot(aes(x=sympt, fill=anxiety)) +
  geom_bar(position="fill")
data %>% 
  ggplot(aes(x=sympt, fill=depress)) +
  geom_bar(position="fill")
```

> Severity of covid symptoms seems to be positively associated with anxiety and depression.

```{r, message=FALSE}
data %>% 
  ggplot(aes(x=anxiety,y=AGE)) +
  geom_boxplot()
data %>% 
  ggplot(aes(x=depress,y=AGE)) +
  geom_boxplot()
```

> Participants with anxiety are on average younger, while participants who feel depressed daily are older between 40 and 60.

```{r, message=FALSE}
data %>% 
  ggplot(aes(x=anxiety,y=MOD10DMIN)) +
  geom_boxplot()
data %>% 
  ggplot(aes(x=depress,y=MOD10DMIN)) +
  geom_boxplot()
```

> Exercise seems to be a protective factor.

```{r, message=FALSE}
data %>% 
  ggplot(aes(x = AGE)) +
  geom_density()
data %>% 
  ggplot(aes(x = MOD10DMIN)) +
  geom_density()
```

# Modeling

## Anxiety - logistic

```{r, message=FALSE}
data_logis <- data %>% dplyr::select(-depress)


logis1 <- glm(anxiety ~ AGE + sexcat + racecat + MOD10DMIN + healthcat + sympt + reg + esupport + incomecat,
              family = binomial, data = data_logis)
logis2 <- glm(anxiety ~ AGE + sexcat + racecat + MOD10DMIN + healthcat + sympt + esupport,
              family = binomial, data = data_logis)
logis3 <- glm(anxiety ~ AGE + sexcat + racecat + healthcat + sympt + esupport,
              family = binomial, data = data_logis)
logis4 <- glm(anxiety ~ AGE + sexcat + racecat + MOD10DMIN + healthcat + sympt,
              family = binomial, data = data_logis)
logis5 <- glm(anxiety ~ AGE + MOD10DMIN + healthcat + sympt + esupport,
              family = binomial, data = data_logis)
logis6 <- glm(anxiety ~ AGE + sexcat + racecat + healthcat + sympt,
              family = binomial, data = data_logis)
logis7 <- glm(anxiety ~ AGE + healthcat + sympt,
              family = binomial, data = data_logis)
logis8 <- glm(anxiety ~ sexcat + racecat + healthcat + sympt + esupport,
              family = binomial, data = data_logis)
logis9 <- glm(anxiety ~ AGE + racecat + healthcat + sympt + esupport,
              family = binomial, data = data_logis)
logis10 <- glm(anxiety ~ AGE + sexcat + healthcat + sympt + esupport,
              family = binomial, data = data_logis)
logis11 <- glm(anxiety ~ AGE + sexcat + racecat + sympt + esupport,
              family = binomial, data = data_logis)
```

```{r}
logis6$coefficients[7:9]

# logis1$coefficients[8:10]
# logis2$coefficients[8:10]
# logis3$coefficients[7:9]
# logis4$coefficients[8:10]
# logis5$coefficients[6:8]
# logis6$coefficients[7:9]
# logis7$coefficients[5:7]
# logis8$coefficients[6:8]
# logis9$coefficients[6:8]
# logis10$coefficients[6:8]
# logis11$coefficients[5:7]

# # test MOD10DMIN
# (logis3$coefficients[7]-logis2$coefficients[8])/logis2$coefficients[8]*100
# (logis3$coefficients[8]-logis2$coefficients[9])/logis2$coefficients[9]*100
# (logis3$coefficients[9]-logis2$coefficients[10])/logis2$coefficients[10]*100

# # test esupport - >10%
# (logis6$coefficients[7]-logis3$coefficients[7])/logis3$coefficients[7]*100
# (logis6$coefficients[8]-logis3$coefficients[8])/logis3$coefficients[8]*100
# (logis6$coefficients[9]-logis3$coefficients[9])/logis3$coefficients[9]*100

# test AGE - >10%
(logis8$coefficients[6]-logis3$coefficients[7])/logis3$coefficients[7]*100
(logis8$coefficients[7]-logis3$coefficients[8])/logis3$coefficients[8]*100
(logis8$coefficients[8]-logis3$coefficients[9])/logis3$coefficients[9]*100

# test sexcat - >10%
(logis9$coefficients[6]-logis3$coefficients[7])/logis3$coefficients[7]*100
(logis9$coefficients[7]-logis3$coefficients[8])/logis3$coefficients[8]*100
(logis9$coefficients[8]-logis3$coefficients[9])/logis3$coefficients[9]*100

# test racecat - >10%
(logis10$coefficients[6]-logis3$coefficients[7])/logis3$coefficients[7]*100
(logis10$coefficients[7]-logis3$coefficients[8])/logis3$coefficients[8]*100
(logis10$coefficients[8]-logis3$coefficients[9])/logis3$coefficients[9]*100

# test healthcat - >10%
(logis11$coefficients[5]-logis3$coefficients[7])/logis3$coefficients[7]*100
(logis11$coefficients[6]-logis3$coefficients[8])/logis3$coefficients[8]*100
(logis11$coefficients[7]-logis3$coefficients[9])/logis3$coefficients[9]*100
```

```{r}
anova(logis2,logis1,test="Chisq")$`Pr(>Chi)` # p>0.05
anova(logis3,logis2,test="Chisq")$`Pr(>Chi)` # p>0.05
anova(logis4,logis2,test="Chisq")$`Pr(>Chi)` # p>0.05
anova(logis5,logis2,test="Chisq")$`Pr(>Chi)` # p<<0.05

anova(logis6,logis3,test="Chisq")$`Pr(>Chi)` # p>0.05
anova(logis6,logis4,test="Chisq")$`Pr(>Chi)` # p>0.05
anova(logis7,logis6,test="Chisq")$`Pr(>Chi)` # p<<0.05

# logis6 based on LRT
```

```{r}
# no significant interaction

logis6_h <- glm(anxiety ~ AGE + sexcat + racecat + healthcat:sympt,
              family = binomial, data = data_logis)
logis6_r <- glm(anxiety ~ AGE + sexcat + racecat:sympt + healthcat,
              family = binomial, data = data_logis)
logis6_s <- glm(anxiety ~ AGE + sexcat:sympt + racecat + healthcat,
              family = binomial, data = data_logis)

anova(logis6,logis6_h,test="Chisq")$`Pr(>Chi)` # p>0.05
anova(logis6,logis6_r,test="Chisq")$`Pr(>Chi)` # p>0.05
anova(logis6,logis6_s,test="Chisq")$`Pr(>Chi)` # p>0.05
```

```{r}
# options(scipen = 999)
tlogis1 = tidy(logis1)
tlogis2 = tidy(logis2)
tlogis3 = tidy(logis3)
tlogis4 = tidy(logis4)
tlogis5 = tidy(logis5)
tlogis6 = tidy(logis6)
tlogis7 = tidy(logis7)

tlogis1$term[tlogis1$p.value < 0.05]
tlogis2$term[tlogis2$p.value < 0.05]
tlogis3$term[tlogis3$p.value < 0.05]
tlogis4$term[tlogis4$p.value < 0.05]
tlogis5$term[tlogis5$p.value < 0.05]
tlogis6$term[tlogis6$p.value < 0.05]
tlogis7$term[tlogis7$p.value < 0.05]

# logis2, logis3 based on number of significant covariates
```

```{r}
# hist(logis$fitted, main="p-hats")
# hist(hatvalues(logis), main="hat-values")

# H0: fit of the model is adequate
options(digits=7)
hoslem.test(logis1$y,fitted(logis1),g=10)
hoslem.test(logis2$y,fitted(logis2),g=10)
hoslem.test(logis3$y,fitted(logis3),g=10)
hoslem.test(logis4$y,fitted(logis4),g=10)
hoslem.test(logis5$y,fitted(logis5),g=10)
hoslem.test(logis6$y,fitted(logis6),g=10)
hoslem.test(logis7$y,fitted(logis7),g=10)

# logis1, logis5, logis7, logis3, logis6, logis2 based on H-L GOF
```

```{r, message=FALSE}
predprob1 <- predict(logis1,type=c("response"))
roccurve1 <- roc(data_logis$anxiety ~ predprob1)
predprob2 <- predict(logis2,type=c("response"))
roccurve2 <- roc(data_logis$anxiety ~ predprob2)
predprob3 <- predict(logis3,type=c("response"))
roccurve3 <- roc(data_logis$anxiety ~ predprob3)
predprob4 <- predict(logis4,type=c("response"))
roccurve4 <- roc(data_logis$anxiety ~ predprob4)
predprob5 <- predict(logis5,type=c("response"))
roccurve5 <- roc(data_logis$anxiety ~ predprob5)
predprob6 <- predict(logis6,type=c("response"))
roccurve6 <- roc(data_logis$anxiety ~ predprob6)
predprob7 <- predict(logis7,type=c("response"))
roccurve7 <- roc(data_logis$anxiety ~ predprob7)
```

```{r}
data.frame(MODEL=c("logis1","logis2","logis3","logis4","logis5","logis6","logis7"),
           AIC=c(AIC(logis1),AIC(logis2),AIC(logis3),AIC(logis4),AIC(logis5),AIC(logis6),AIC(logis7)),
           BIC=c(BIC(logis1),BIC(logis2),BIC(logis3),BIC(logis4),BIC(logis5),BIC(logis6),BIC(logis7)),
           AUC=c(auc(roccurve1),auc(roccurve2),auc(roccurve3),auc(roccurve4),auc(roccurve5),auc(roccurve6),auc(roccurve7)))
# logis3, logis6 based on AIC
# logis7, logis6 based on BIC
# logis1, logis3, logis2, logis6 based on AUC

# stepModel <- step(logis1, direction=c("both"))
# sexcat + racecat + healthcat + esupport based on stepwise selection

pred<-ifelse(predict(logis3,type="response")>0.5,"Yes","No")
actu<-data_logis$anxiety
confusionMatrix(table(pred,actu))

pred<-ifelse(predict(logis6,type="response")>0.5,"Yes","No")
actu<-data_logis$anxiety
confusionMatrix(table(pred,actu))

VIF(logis1) # no collinearity
VIF(logis3) # no collinearity
VIF(logis6) # no collinearity
```

## Depression - multinomial & ordinal

```{r, message=FALSE, results='hide'}
data_multi <- data %>% dplyr::select(-anxiety)

multi1 <- multinom(depress ~ AGE + sexcat + racecat + MOD10DMIN + healthcat + sympt + reg + esupport + incomecat, data = data_multi)
multi2 <- multinom(depress ~ AGE + sexcat + racecat + MOD10DMIN + healthcat + sympt + esupport, data = data_multi)
multi3 <- multinom(depress ~ AGE + sexcat + racecat + healthcat + sympt + esupport, data = data_multi)
multi4 <- multinom(depress ~ AGE + sexcat + racecat + MOD10DMIN + healthcat + sympt, data = data_multi)
multi5 <- multinom(depress ~ AGE + MOD10DMIN + healthcat + sympt + esupport, data = data_multi)
multi6 <- multinom(depress ~ AGE + sexcat + racecat + healthcat + sympt, data = data_multi)
multi7 <- multinom(depress ~ AGE + healthcat + sympt, data = data_multi)
multi8 <- multinom(depress ~ sexcat + racecat + healthcat + sympt, data = data_multi)
multi9 <- multinom(depress ~ AGE + racecat + healthcat + sympt, data = data_multi)
multi10 <- multinom(depress ~ AGE + sexcat + healthcat + sympt, data = data_multi)
multi11 <- multinom(depress ~ AGE + sexcat + racecat + sympt, data = data_multi)
```

```{r}
coef(multi1)[,8:10]
coef(multi2)[,8:10]
coef(multi3)[,7:9]
coef(multi4)[,8:10]
coef(multi5)[,6:8]
coef(multi6)[,7:9]
coef(multi7)[,5:7]
coef(multi8)[,6:8]
coef(multi9)[,6:8]
coef(multi10)[,6:8]
coef(multi11)[,5:7]
```

```{r}
# # test MOD10DMIN
# (coef(multi3)[1,7]-coef(multi2)[1,8])/coef(multi2)[1,8]*100
# (coef(multi3)[1,8]-coef(multi2)[1,9])/coef(multi2)[1,9]*100
# (coef(multi3)[1,9]-coef(multi2)[1,10])/coef(multi2)[1,10]*100
# 
# (coef(multi3)[2,7]-coef(multi2)[2,8])/coef(multi2)[2,8]*100
# (coef(multi3)[2,8]-coef(multi2)[2,9])/coef(multi2)[2,9]*100
# (coef(multi3)[2,9]-coef(multi2)[2,10])/coef(multi2)[2,10]*100
# 
# (coef(multi3)[3,7]-coef(multi2)[3,8])/coef(multi2)[3,8]*100
# (coef(multi3)[3,8]-coef(multi2)[3,9])/coef(multi2)[3,9]*100
# (coef(multi3)[3,9]-coef(multi2)[3,10])/coef(multi2)[3,10]*100
```

```{r}
# test AGE - >10%
(coef(multi8)[1,6]-coef(multi6)[1,7])/coef(multi6)[1,7]*100
(coef(multi8)[1,7]-coef(multi6)[1,8])/coef(multi6)[1,8]*100
(coef(multi8)[1,8]-coef(multi6)[1,9])/coef(multi6)[1,9]*100

(coef(multi8)[2,6]-coef(multi6)[2,7])/coef(multi6)[2,7]*100
(coef(multi8)[2,7]-coef(multi6)[2,8])/coef(multi6)[2,8]*100
(coef(multi8)[2,8]-coef(multi6)[2,9])/coef(multi6)[2,9]*100

(coef(multi8)[3,6]-coef(multi6)[3,7])/coef(multi6)[3,7]*100
(coef(multi8)[3,7]-coef(multi6)[3,8])/coef(multi6)[3,8]*100
(coef(multi8)[3,8]-coef(multi6)[3,9])/coef(multi6)[3,9]*100
```

```{r}
# # test sexcat
# (coef(multi9)[1,6]-coef(multi6)[1,7])/coef(multi6)[1,7]*100
# (coef(multi9)[1,7]-coef(multi6)[1,8])/coef(multi6)[1,8]*100
# (coef(multi9)[1,8]-coef(multi6)[1,9])/coef(multi6)[1,9]*100
# 
# (coef(multi9)[2,6]-coef(multi6)[2,7])/coef(multi6)[2,7]*100
# (coef(multi9)[2,7]-coef(multi6)[2,8])/coef(multi6)[2,8]*100
# (coef(multi9)[2,8]-coef(multi6)[2,9])/coef(multi6)[2,9]*100
# 
# (coef(multi9)[3,6]-coef(multi6)[3,7])/coef(multi6)[3,7]*100
# (coef(multi9)[3,7]-coef(multi6)[3,8])/coef(multi6)[3,8]*100
# (coef(multi9)[3,8]-coef(multi6)[3,9])/coef(multi6)[3,9]*100
```

```{r}
# test racecat - >10%
(coef(multi10)[1,6]-coef(multi6)[1,7])/coef(multi6)[1,7]*100
(coef(multi10)[1,7]-coef(multi6)[1,8])/coef(multi6)[1,8]*100
(coef(multi10)[1,8]-coef(multi6)[1,9])/coef(multi6)[1,9]*100

(coef(multi10)[2,6]-coef(multi6)[2,7])/coef(multi6)[2,7]*100
(coef(multi10)[2,7]-coef(multi6)[2,8])/coef(multi6)[2,8]*100
(coef(multi10)[2,8]-coef(multi6)[2,9])/coef(multi6)[2,9]*100

(coef(multi10)[3,6]-coef(multi6)[3,7])/coef(multi6)[3,7]*100
(coef(multi10)[3,7]-coef(multi6)[3,8])/coef(multi6)[3,8]*100
(coef(multi10)[3,8]-coef(multi6)[3,9])/coef(multi6)[3,9]*100
```

```{r}
# test healthcat - >10%
(coef(multi11)[1,5]-coef(multi6)[1,7])/coef(multi6)[1,7]*100
(coef(multi11)[1,6]-coef(multi6)[1,8])/coef(multi6)[1,8]*100
(coef(multi11)[1,7]-coef(multi6)[1,9])/coef(multi6)[1,9]*100

(coef(multi11)[2,5]-coef(multi6)[2,7])/coef(multi6)[2,7]*100
(coef(multi11)[2,6]-coef(multi6)[2,8])/coef(multi6)[2,8]*100
(coef(multi11)[2,7]-coef(multi6)[2,9])/coef(multi6)[2,9]*100

(coef(multi11)[3,5]-coef(multi6)[3,7])/coef(multi6)[3,7]*100
(coef(multi11)[3,6]-coef(multi6)[3,8])/coef(multi6)[3,8]*100
(coef(multi11)[3,7]-coef(multi6)[3,9])/coef(multi6)[3,9]*100
```

```{r}
anova(multi2,multi1,test="Chisq")$`Pr(Chi)` # p>0.05
anova(multi3,multi2,test="Chisq")$`Pr(Chi)` # p>0.05
anova(multi4,multi2,test="Chisq")$`Pr(Chi)` # p<<0.05
anova(multi5,multi3,test="Chisq")$`Pr(Chi)` # p>0.05

anova(multi6,multi5,test="Chisq")$`Pr(Chi)` # p<<0.05
anova(multi7,multi5,test="Chisq")$`Pr(Chi)` # p<<0.05
anova(multi9,multi5,test="Chisq")$`Pr(Chi)` # p<<0.05

# multi5 based on LRT
```

```{r, message=FALSE, results='hide'}
# no significant interaction

multi6_r <- multinom(depress ~ AGE + sexcat + racecat:sympt + healthcat, data = data_multi)
multi6_h <- multinom(depress ~ AGE + sexcat + racecat + healthcat:sympt, data = data_multi)

anova(multi6,multi6_h,test="Chisq")$`Pr(Chi)` # p>0.05
anova(multi6,multi6_r,test="Chisq")$`Pr(Chi)` # p>0.05
```

```{r}
data.frame(MODEL=c("multi1","multi2","multi3","multi4","multi5","multi6","multi7"),
           AIC=c(AIC(multi1),AIC(multi2),AIC(multi3),AIC(multi4),AIC(multi5),AIC(multi6),AIC(multi7)),
           BIC=c(BIC(multi1),BIC(multi2),BIC(multi3),BIC(multi4),BIC(multi5),BIC(multi6),BIC(multi7)))

# multi5, multi3 based on AIC
# multi7, multi5 based on BIC
```

```{r}
logitgof(data_multi$depress, fitted(multi1))
logitgof(data_multi$depress, fitted(multi2))
logitgof(data_multi$depress, fitted(multi3))
logitgof(data_multi$depress, fitted(multi4))
logitgof(data_multi$depress, fitted(multi5))
logitgof(data_multi$depress, fitted(multi6))
logitgof(data_multi$depress, fitted(multi7))

# all p>0.05
```

```{r}
# options(scipen = 999)
tmulti1 = tidy(multi1)
tmulti2 = tidy(multi2)
tmulti3 = tidy(multi3)
tmulti4 = tidy(multi4)
tmulti5 = tidy(multi5)
tmulti6 = tidy(multi6)
tmulti7 = tidy(multi7)

# tmulti1$term[tmulti1$p.value < 0.05]
# tmulti2$term[tmulti2$p.value < 0.05]
# tmulti3$term[tmulti3$p.value < 0.05]
# tmulti4$term[tmulti4$p.value < 0.05]
# tmulti5$term[tmulti5$p.value < 0.05]
# tmulti7$term[tmulti7$p.value < 0.05]
tmulti6$term[tmulti6$p.value < 0.05]

# multi1 based on number of significant covariates
```

```{r}
# fitted probs close to actual probs

cbind(fitted(multi5)[data_multi$sympt == "no symptoms",][1,],fitted(multi5)[data_multi$sympt == "mild symptoms",][1,],fitted(multi5)[data_multi$sympt == "moderate symptoms",][1,],fitted(multi5)[data_multi$sympt == "severe symptoms",][1,])
prop.table(table(data_multi$depress,data_multi$sympt),2)
```

```{r}
# never
plot(na.omit(data_multi$sympt), fitted(multi5)[,1])

# daily
plot(na.omit(data_multi$sympt), fitted(multi5)[,2], main="Figure 1. Predicted Probabilities of Feeling Depressed Daily", xlab="COVID-19 Symptom", ylab="Probability")

# weekly or monthly
plot(na.omit(data_multi$sympt), fitted(multi5)[,3])

# yearly
plot(na.omit(data_multi$sympt), fitted(multi5)[,4])
```

```{r, eval=FALSE}
# proportional assumption does not hold

data_multi$depressdaily <- ifelse(data_multi$depress=="daily", 1, 0)
data_multi$depressnever <- ifelse(data_multi$depress=="never", 1, 0)
moddaily <- glm(depressdaily ~ ., family=binomial, data=data_multi %>% dplyr::select(-depress, -depressnever)) 
summary(moddaily)
modnever <- glm(depressnever ~ ., family=binomial, data=data_multi %>% dplyr::select(-depress, -depressdaily)) 
summary(modnever)
coef(moddaily)[10:12]
coef(moddaily)[10] + c(-1, 1)*1.96*sqrt(vcov(moddaily)[10,10])
coef(moddaily)[11] + c(-1, 1)*1.96*sqrt(vcov(moddaily)[11,11])
coef(moddaily)[12] + c(-1, 1)*1.96*sqrt(vcov(moddaily)[12,12])
coef(modnever)[10:12]
coef(modnever)[10] + c(-1, 1)*1.96*sqrt(vcov(modnever)[10,10])
coef(modnever)[11] + c(-1, 1)*1.96*sqrt(vcov(modnever)[11,11])
coef(modnever)[12] + c(-1, 1)*1.96*sqrt(vcov(modnever)[12,12])
```

# Kenny's EDA

```{r}
# univariate
hist(data$MOD10DMIN)
# tally(~data$MOD10FWK)

hist(log(data$MOD10DMIN))
hist(data$AGE)
barplot(prop.table(table(data$sympt)))
barplot(prop.table(table(data$sexcat)))
barplot(prop.table(table(data$racecat)))
barplot(prop.table(table(data$incomecat)))
barplot(prop.table(table(data$healthcat)))
ggplot(data, aes(sympt, MOD10DMIN)) +   
  geom_boxplot() +
  labs(x = "Severity of COVID-19 Symptoms", y = "Duration of Moderate Physical Activity")
```

We see that the duration of moderate physical activity is heavily right skewed so we log transformed it.

```{r}
#bivariate
scatter.smooth(data$AGE, log(data$MOD10DMIN))
scatter.smooth(data$AGE, (data$MOD10DMIN))
boxplot(data$MOD10DMIN ~ data$sympt)
boxplot(data$MOD10DMIN ~ data$racecat)
boxplot(data$MOD10DMIN ~ data$incomecat)
boxplot(data$MOD10DMIN ~ data$healthcat)
```

# Fitting the full model

```{r}
lm1 <- lm(log(MOD10DMIN) ~ sexcat + racecat  + healthcat + sympt + AGE
      ,data = data)

summary(lm1)
```

In the full model, we see that after adjusting for the effect of the other variables, none of the COVID symptom levels have a statistically significant effect on the duration of moderate physical activity.

# Effect modifier

```{r}
# check if anxiety, depression, region or emotional support are effect modifiers of covid symptoms
lm_anxiety <- lm(log(MOD10DMIN) ~ sexcat + racecat  + healthcat + sympt + AGE
          + anxiety*sympt,data = data)
lm_depress <- lm(log(MOD10DMIN) ~ sexcat + racecat  + healthcat + sympt + AGE
          + depress*sympt,data = data)
lm_reg <- lm(log(MOD10DMIN) ~ sexcat + racecat  + healthcat + sympt + AGE
          + reg*sympt,data = data)
lm_esupport <- lm(log(MOD10DMIN) ~ sexcat + racecat + incomecat + healthcat + sympt + AGE
          + esupport*sympt,data = data)
lm_income <- lm(log(MOD10DMIN) ~ sexcat + racecat  + healthcat + sympt + AGE
          + incomecat*sympt,data = data)

summary(lm_anxiety)
summary(lm_depress)
summary(lm_reg)
summary(lm_esupport)
summary(lm_income)
# None are effect modifiers so we will stick with
lm1 <- lm(log(MOD10DMIN) ~ sexcat + racecat + healthcat + sympt + AGE, data = data)

summary(lm1)
```

# Diagnosis and Model Evaluation

```{r}
plot(lm1, 1:6)
hist(lm1$residuals)
# 265, 46, 383

# 12/397
# 0.0302267

# modelling dropping those 
lm1_outliers <- lm(log(MOD10DMIN) ~ sexcat + racecat + healthcat + sympt + AGE, data = 
data[-c(46,265, 383),])

summary(lm1_outliers)
summary(lm1)

# none are above 5 :)
VIF(lm1) 
```

From the residual plots, it seems linearity, normality, and equal variance holds.
